---
title: "DA 401 Preliminary Results"
author: "Elysia Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATA PREPROCESSING

### Import individual datasets

```{r}
ExchangeRate_Annual <- read.csv("AnnualExchangeRate.csv", header = TRUE)
ProductionValue <- read.csv("ProductionValue.csv", header = TRUE)
ProductionQuantity <- read.csv("ProductionQuantity.csv", header = TRUE)
ExportQuantity <- read.csv("ExportQuantity.csv", header = TRUE)
```

Export Quantity: FAOSTAT has missing data for “Rice.” “Rice, milled” is used and treated as equivalent to “Rice” in others.

Production Value: https://files-faostat.fao.org/production/QV/QV_e.pdf. Understand constant units and what can or cannot be inferred from them. 

Producer Price: https://files-faostat.fao.org/production/QV/QV_e.pdf. Justify the calculation. Note the unit.

Determine the common headers or columns to include in the overall dataset:
- Element for variable name
- Year
- Unit (not included in the dataset but noted in the metadata)
- Value

### Extract relevant columns from each datasets

```{r}
library(dplyr)

Rice_ExchangeRate_Annual <- ExchangeRate_Annual %>% select(Element, Year, Value)
Rice_ProductionValue <- ProductionValue %>% select(Element, Year, Value)
Rice_ProductionQuantity <- ProductionQuantity %>% select(Element, Year, Value)
Rice_ExportQuantity <- ExportQuantity %>% select(Element, Year, Value)
```

### Combine individual datasets into one overall 

```{r}
library(dplyr)

Rice_Unpivot <- bind_rows(
  Rice_ExchangeRate_Annual, 
  Rice_ProductionValue, 
  Rice_ProductionQuantity, 
  Rice_ExportQuantity) %>%
  arrange(Year) # Arrange rows by year

head(Rice_Unpivot)
```

### Pivot the dataset with each variable as its own column

```{r}
library(tidyr)

Rice <- Rice_Unpivot %>% pivot_wider(names_from = Element, values_from = Value)

head(Rice)
```

### Rename the variables

```{r}
Rice <- Rice %>%
  rename("VNDPerUSD" = "Standard local currency units per USD",
         "ProductionValueVND" = "Gross Production Value (constant 2014-2016 thousand SLC)",
         "ProductionValueUSD" = "Gross Production Value (constant 2014-2016 thousand US$)",
         "ProductionQuantity" = "Production",
         "ExportQuantity" = "Export quantity")

head(Rice)
```

### Calculate Producer Price as Production Value divided by Production Quantity

```{r}
Rice <- Rice %>% mutate(ProducerPriceVND = ProductionValueVND / ProductionQuantity, 
                        ProducerPriceUSD = ProductionValueUSD / ProductionQuantity)

head(Rice)
```

Note the units in the documentation.

*Values of Producer Prices are identical throughout the years. Use Production Values in VAR instead? Production Values are not standardized.* 

# DATA VISUALIZATION

### Visualize the variables over time highlighting the devaluation in 1985

```{r}
library(ggplot2)

ggplot(Rice, aes(x = Year, y = VNDPerUSD)) +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "red") +
  labs(title = "Vietnam Dong (VND) Per United States Dollar (USD) Exchange Rates 1975-2023",
       x = "Year",
       y = "VND Per USD Exchange Rate") +
  theme_classic()
```

```{r}
ggplot(Rice, aes(x = Year, y = ProducerPriceUSD)) +
  geom_line(color = "green") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "red") +
  labs(title = "Producer Prices of Rice in United States Dollars (USD) 1975-2023",
       x = "Year",
       y = "Price in USD") +
  theme_classic()
```

```{r}
ggplot(Rice, aes(x = Year, y = ProducerPriceVND)) +
  geom_line(color = "darkgreen") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "red") +
  labs(title = "Producer Prices of Rice in Vietnam Dong 1975-2023",
       x = "Year",
       y = "Price in VND") +
  theme_classic()
```

```{r}
ggplot(Rice, aes(x = Year, y = ProductionValueUSD)) +
  geom_line(color = "green") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "red") +
  labs(title = "Production Values of Rice in United States Dollars (USD) 1975-2023",
       x = "Year",
       y = "Value in USD") +
  theme_classic()
```

```{r}
ggplot(Rice, aes(x = Year, y = ProductionValueVND)) +
  geom_line(color = "darkgreen") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "red") +
  labs(title = "Production Values of Rice in Vietnam Dong (VND) 1975-2023",
       x = "Year",
       y = "Value in VND") +
  theme_classic()
```

```{r}
ggplot(Rice, aes(x = Year, y = ExportQuantity)) +
  geom_line(color = "purple") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "red") +
  labs(title = "Export Quantities of Rice in Tons 1975-2023",
       x = "Year",
       y = "Export in Tons") +
  theme_classic()
```

Looking at the y-axis, I should consider transforming the values of Export Quantity.

# VAR MODEL

### Make each variable a time series

*I have yet to control for GDP or inflation.*

```{r}
ExchangeRate_Annual_TS <- ts(Rice$VNDPerUSD, 
                              start = min(Rice$Year), 
                              end = max(Rice$Year), 
                              frequency = 1)

ProducerPriceVND_TS <- ts(Rice$ProducerPriceVND, start = min(Rice$Year), end = max(Rice$Year), frequency = 1)

ProducerPriceUSD_TS <- ts(Rice$ProducerPriceUSD, start = min(Rice$Year), end = max(Rice$Year), frequency = 1)

ExportQuantity_TS <- ts(Rice$ExportQuantity, 
                         start = min(Rice$Year), 
                         end = max(Rice$Year), 
                         frequency = 1)

# Use Production Values instead of Producer Prices

ProductionValueVND_TS <- ts(Rice$ProductionValueVND, 
                           start = min(Rice$Year), 
                           end = max(Rice$Year), 
                           frequency = 1)

ProductionValueUSD_TS <- ts(Rice$ProductionValueUSD, 
                           start = min(Rice$Year), 
                           end = max(Rice$Year), 
                           frequency = 1)
```

There is no missing data for each variable or time series.

### Examine stationarity of each time series

```{r}
plot(ExchangeRate_Annual_TS, 
     main = "Vietnam Dong (VND) Per United States Dollar (USD) Exchange Rates 1975-2023",
     xlab = "Year", ylab = "VND Per USD Exchange Rate")
```

There is likely no need to make variations of the same line charts.

```{r}
acf(ExchangeRate_Annual_TS, main = "ACF of Exchange Rates")
```

```{r}
pacf(ExchangeRate_Annual_TS, main = "PACF of Exchange Rates")
```

```{r}
library(tseries)

adf.test(ExchangeRate_Annual_TS)

adf.test(ExportQuantity_TS)

adf.test(ProductionValueUSD_TS)

adf.test(ProductionValueVND_TS)
```

None of the time series is stationary.

### Transform each time series into a stationary time series through differencing

```{r}
ExchangeRate_Annual_Diff <- diff(ExchangeRate_Annual_TS, differences = 1)

adf.test(ExchangeRate_Annual_Diff)
```

```{r}
#ProducerPriceVND_Diff <- diff(ProducerPriceVND_TS, differences = 2)

#adf.test(ProducerPriceVND_Diff)
```

I need to take the second difference of all the time series because one difference is insufficient to make ProducerPriceVND_TS stationary.

```{r}
#ProducerPriceUSD_Diff <- diff(ProducerPriceUSD_TS, differences = 2)

#adf.test(ProducerPriceUSD_Diff)
```

```{r}
ExportQuantity_Diff <- diff(ExportQuantity_TS, differences = 1)

adf.test(ExportQuantity_Diff)
```

```{r}
ProductionValueVND_Diff <- diff(ProductionValueVND_TS, differences = 1)

adf.test(ExportQuantity_Diff)
```

```{r}
ProductionValueUSD_Diff <- diff(ProductionValueUSD_TS, differences = 1)

adf.test(ExportQuantity_Diff)
```

*What is cointegration?*

### Combine all the time series into one matrix

```{r}
Rice_Matrix <- cbind(ExchangeRate_Annual_Diff, ProductionValueUSD_Diff, ExportQuantity_Diff) 
```

Production Values in USD is used in the model as they are expected to reflect the fluctuations in Exchange Rates more clearly than those in VND.

### Determine the lag order

```{r}
library(vars)

VARselect(Rice_Matrix, lag.max = 3, type = "none")
```

*const, trend, both?*

*How to read the results? Do they indicate 2 overall?*

### Estimate the VAR model

```{r}
#nrow(Rice_Matrix)
```

```{r}
#cor(Rice_Matrix)
```

```{r}
#summary(Rice_Matrix)
```

```{r}
Rice_VAR <- VAR(Rice_Matrix, p = 2, type = "none")

summary(Rice_VAR)
```

*Error in solve.default(Sigma): system is computationally singular: reciprocal condition number = 2.05396e-28* when Producer Prices are used.

# Perform Cholesky Decomposition

```{r}
# Extract residuals from the estimated VAR model
Cholesky_resid <- resid(Rice_VAR)

# Compute the covariance matrix of the residuals
Cholesky_resid_cov <- cov(Cholesky_resid)

# Perform Cholesky decomposition
# chol() returns an upper-triangular matrix, so transpose it to obtain a lower-triangular matrix
Cholesky_decomp <- t(chol(Cholesky_resid_cov))

# Check the decomposition
Cholesky_decomp %*% t(Cholesky_decomp)
```

*What does any of this mean?*

# Compute IRF

```{r}
IRF_general <- irf(Rice_VAR, impulse = "ExchangeRate_Annual_Diff", response = c("ProductionValueUSD_Diff", "ExportQuantity_Diff"), n.ahead = 10, ortho = FALSE, boot = TRUE)

plot(IRF_general)
```

