---
title: "DA 401 Paper Draft"
author: "Elysia Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATA PREPROCESSING

*Import and process monthly exchange rates.*

### Import individual datasets

```{r}
ExchangeRate_Annual <- read.csv("AnnualExchangeRate.csv", header = TRUE)
ExportQuantity <- read.csv("ExportQuantity.csv", header = TRUE)
```

Export Quantity: FAOSTAT has missing data for “Rice.” “Rice, milled” is used and treated as equivalent to “Rice” in others.

Determine the common headers or columns to include in the overall dataset:
- Element for variable name
- Year
- Unit (not included in the dataset but noted in the metadata)
- Value

```{r}
ExchangeRate_Monthly <- read.csv("MonthlyExchangeRate.csv", header = TRUE)
```

### Extract relevant columns from each datasets

```{r}
library(dplyr)

Rice_ExchangeRate_Annual <- ExchangeRate_Annual %>% select(Element, Year, Value)
Rice_ExportQuantity <- ExportQuantity %>% select(Element, Year, Value)
```

```{r}
ExchangeRate_Monthly <- ExchangeRate_Monthly %>% select(Year, Months, Value) 

MonthLevels <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

ExchangeRate_Monthly$Months <- factor(ExchangeRate_Monthly$Months, levels = MonthLevels, ordered = TRUE)

ExchangeRate_Monthly <- ExchangeRate_Monthly %>% arrange(Year, Months)

```

### Combine individual datasets into one overall 

```{r}
Rice_Unpivot <- bind_rows(
  Rice_ExchangeRate_Annual, 
  Rice_ExportQuantity) %>%
  arrange(Year) # Arrange rows by year

head(Rice_Unpivot)
```

### Pivot the dataset with each variable as its own column

```{r}
library(tidyr)

Rice <- Rice_Unpivot %>% pivot_wider(names_from = Element, values_from = Value)

head(Rice)
```

### Rename the variables

```{r}
Rice <- Rice %>%
  rename("VNDPerUSD" = "Standard local currency units per USD",
         "ExportQuantity" = "Export quantity")

head(Rice)
```

Note the units in the documentation.

# DATA VISUALIZATION

*Analyze monthly exchange rates.*

### Visualize the variables over time highlighting the devaluation in 1985

```{r}
library(ggplot2)
library(scales)

ggplot(Rice, aes(x = Year, y = VNDPerUSD)) +
  geom_line(size = 0.75, color = "steelblue") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "firebrick") +
  labs(title = "Vietnam Dong (VND) Per United States Dollar (USD) Exchange Rates 1975-2023",
       x = "Year",
       y = "VND Per USD Exchange Rate") +
  theme_classic() +
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(face = "bold")) +
  scale_y_continuous(labels = comma)
```

```{r}
ggplot(Rice, aes(x = Year, y = ExportQuantity)) +
  geom_line(size = 0.75, color = "forestgreen") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "firebrick") +
  labs(title = "Export Quantities of Rice in Tons 1975-2023",
       x = "Year",
       y = "Export in Tons") +
  theme_classic() +
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(face = "bold")) +
  scale_y_continuous(labels = comma)
```

Looking at the y-axis, I should consider transforming the values of Export Quantity.

```{r}
ggplot(ExchangeRate_Monthly, aes(x = Months, y = Value, group = Year)) +
  geom_line(size = 0.75, color = "firebrick") +
  labs(title = "Vietnam Dong (VND) Per United States Dollar (USD) Exchange Rates 1985",
       x = "Month",
       y = "VND Per USD Exchange Rate") +
  theme_classic() +
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(face = "bold"))
```

# SVAR MODEL

### Make each variable a time series

*I have yet to control for GDP or inflation.*

```{r}
ExchangeRate_Annual_TS <- ts(Rice$VNDPerUSD, 
                              start = min(Rice$Year), 
                              end = max(Rice$Year), 
                              frequency = 1)

ExportQuantity_TS <- ts(Rice$ExportQuantity, 
                         start = min(Rice$Year), 
                         end = max(Rice$Year), 
                         frequency = 1)
```

There is no missing data for each variable or time series.

### Examine stationarity of each time series

```{r}
acf(ExchangeRate_Annual_TS, main = "ACF of Annual Exchange Rates")
```

```{r}
pacf(ExchangeRate_Annual_TS, main = "PACF of Annual Exchange Rates")
```

```{r}
acf(ExportQuantity_TS, main = "ACF of Export Quantities")
```

```{r}
pacf(ExportQuantity_TS, main = "PACF of Export Quantities")
```

```{r}
library(tseries)

adf.test(ExchangeRate_Annual_TS)

adf.test(ExportQuantity_TS)
```

None of the time series is stationary.

### Transform each time series into a stationary time series through differencing

```{r}
ExchangeRate_Annual_Diff <- diff(ExchangeRate_Annual_TS, differences = 1)

adf.test(ExchangeRate_Annual_Diff)
```

```{r}
ExportQuantity_Diff <- diff(ExportQuantity_TS, differences = 1)

adf.test(ExportQuantity_Diff)
```

### Combine all the time series into one matrix

```{r}
Rice_Matrix <- cbind(ExchangeRate_Annual_Diff, ExportQuantity_Diff) 
```

### Determine the lag order

```{r}
library(vars)

VARselect(Rice_Matrix, lag.max = 3, type = "none")
```

*const, trend, both?*

*How to read the results? Do they indicate 2 overall?*

### Estimate the reduced-form VAR

```{r}
Rice_VAR <- VAR(Rice_Matrix, p = 2, type = "none")

summary(Rice_VAR)
```

### Identify structural shocks with SVAR CV

*Choose a specific method based on my theoretical framework.*

- Heteroskedasticity-based (CV, ST, GARCH): leverage time-varying variances in shocks to identify structural parameters.

=> This method assumes two different types of variance. Try all three (ST) to see which one fits the best. Note that these three make different assumptions themselves (page 15).

=> This might be a safer assumption.

- Independence-based (CVM, DC, NGML): rely on non-Gaussian innovations and statistical independence to identify shocks.

=> This method assumes non-normal distribution of the residuals.

*Model estimation and identification* 

*Diagnostic testing (e.g., Chow tests, Wald tests)*

- B matrix is a unique decomposition of the covariance matrix.

- Wald test guarantees lambda 1 and 2 are different.

=> This distinguishes the two variables.

- VAR u = SVAR B * e

*Inference via bootstrap methods (wild, moving-block, bootstrap-after-bootstrap)*

*Economic interpretation tools like impulse-response functions, FEVD, and historical decomposition*

- Show CF for IRF.

- Historical decomposition and counterfactual analysis come after.

```{r}
library(svars)

Rice_SVAR <- id.cv(Rice_VAR, SB = c(1985, 1)) # Specify the structural break in 1985

summary(Rice_SVAR)
```

### Assess the IRFs

```{r}
plot(irf(Rice_SVAR, n.ahead = 10))
```

### Add confidence intervals around IRFs

```{r}
Rice_IRF <- wild.boot(Rice_SVAR, nboot = 500)
```

```{r}
plot(Rice_IRF)
```

### Compute FEVD

```{r}
fevd(Rice_SVAR, n.ahead = 10)
```

### Compute the historical decomposition

```{r}
Rice_HD <- hd(Rice_SVAR)

plot(Rice_HD)
```

### Conduct counterfactual analysis
