---
title: "From Markets to Fields: The Impact of Vietnam’s 1985 Dong Devaluation on Rice Exports"
author: "Elysia Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATA PREPROCESSING

### Import individual datasets

```{r}
ExchangeRate_Annual <- read.csv("AnnualExchangeRate.csv", header = TRUE)
ExportQuantity <- read.csv("ExportQuantity.csv", header = TRUE)
RealGDP <- read.csv("RealGDP.csv", header = TRUE)
GDPDeflator_VND <- read.csv("VNDGDPDeflator.csv", header = TRUE)
GDPDeflator_USD <- read.csv("USDGDPDeflator.csv", header = TRUE)
ExchangeRate_Monthly <- read.csv("MonthlyExchangeRate.csv", header = TRUE)
```

### Extract relevant columns from each datasets

```{r}
library(dplyr)

Rice_ExchangeRate_Annual <- ExchangeRate_Annual %>% select(Element, Year, Value)
Rice_ExportQuantity <- ExportQuantity %>% select(Element, Year, Value)
Rice_RealGDP <- RealGDP %>% select(Element, Year, Value)
Rice_GDPDeflator_VND <- GDPDeflator_VND %>% select(Element, Year, Value)
Rice_GDPDeflator_USD <- GDPDeflator_USD %>% select(Element, Year, Value)
```

```{r}
ExchangeRate_Monthly <- ExchangeRate_Monthly %>% select(Year, Months, Value) 

MonthLevels <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

ExchangeRate_Monthly$Months <- factor(ExchangeRate_Monthly$Months, levels = MonthLevels, ordered = TRUE)

ExchangeRate_Monthly <- ExchangeRate_Monthly %>% arrange(Year, Months)
```

### Pivot each individual dataset then rename the variable

```{r}
library(tidyr)

Rice_ExchangeRate_Annual <- Rice_ExchangeRate_Annual %>% pivot_wider(names_from = Element, values_from = Value)
Rice_ExchangeRate_Annual <- Rice_ExchangeRate_Annual %>% rename("VNDPerUSD" = "Standard local currency units per USD")
                                                                
Rice_ExportQuantity <- Rice_ExportQuantity %>% pivot_wider(names_from = Element, values_from = Value)
Rice_ExportQuantity <- Rice_ExportQuantity %>% rename("ExportQuantity" = "Export quantity")
                  
Rice_RealGDP <- Rice_RealGDP %>% pivot_wider(names_from = Element, values_from = Value)
Rice_RealGDP <- Rice_RealGDP %>% rename("RealGDP" = "Value Standard Local Currency, 2015 prices")

Rice_GDPDeflator_VND <- Rice_GDPDeflator_VND %>% pivot_wider(names_from = Element, values_from = Value)
Rice_GDPDeflator_VND <- Rice_GDPDeflator_VND %>% rename("VNDGDPDeflator" = "Value Standard Local Currency, 2015 prices")

Rice_GDPDeflator_USD <- Rice_GDPDeflator_USD %>% pivot_wider(names_from = Element, values_from = Value)
Rice_GDPDeflator_USD <- Rice_GDPDeflator_USD %>% rename("USDGDPDeflator" = "Value Standard Local Currency, 2015 prices")
```

### Combine individual datasets into one overall and arrange records by year

```{r}
Rice <- Rice_ExchangeRate_Annual %>%
  left_join(Rice_ExportQuantity, by = "Year") %>%
  left_join(Rice_RealGDP, by = "Year") %>%
  left_join(Rice_GDPDeflator_VND, by = "Year") %>%
  left_join(Rice_GDPDeflator_USD, by = "Year") %>%
  arrange(Year)

head(Rice)
```

### Calculate real exchange rates as nominal exchange rates multiplied by GDP deflators

```{r}
Rice <- Rice %>% mutate(RealVNDPerUSD = VNDPerUSD*(USDGDPDeflator/VNDGDPDeflator))

head(Rice)
```

# DATA VISUALIZATION

### Visualize the variables over time highlighting the devaluation in 1985

*Monthly exchange rates*

```{r}
library(ggplot2)
library(scales)

Plot_ExchangeRate_Monthly <- ggplot(ExchangeRate_Monthly, aes(x = Months, y = Value, group = Year)) +
  geom_line(size = 0.75, color = "firebrick") +
  labs(x = "Month",
       y = "VND Per USD Exchange Rate") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9)  
  )

Plot_ExchangeRate_Monthly
```

*Nominal annual exchange rates*

```{r}
Plot_NER <- ggplot(Rice, aes(x = Year, y = VNDPerUSD)) +
  geom_line(size = 0.75, color = "steelblue") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "firebrick") +
  labs(x = "Year",
       y = "Nominal VND Per USD Exchange Rate") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9)  
  ) +
  scale_y_continuous(labels = comma)

Plot_NER
```

*Real annual exchange rates*

```{r}
Plot_RER <- ggplot(Rice, aes(x = Year, y = RealVNDPerUSD)) +
  geom_line(size = 0.75, color = "deepskyblue4") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "firebrick") +
  labs(x = "Year",
       y = "Real VND Per USD Exchange Rate") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9)  
  ) +
  scale_y_continuous(labels = comma)

Plot_RER
```

*Annual export quantities*

```{r}
Plot_Export <- ggplot(Rice, aes(x = Year, y = ExportQuantity)) +
  geom_line(size = 0.75, color = "darkolivegreen") +
  scale_x_continuous(breaks = seq(min(Rice$Year), max(Rice$Year), by = 5)) +
  geom_vline(xintercept = 1985, linetype = "dashed", color = "firebrick") +
  labs(x = "Year",
       y = "Export in Tons") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9)  
  ) +
  scale_y_continuous(labels = comma)

Plot_Export
```

# SVAR MODEL

### Make each variable a time series

```{r}
ExportQuantity_TS <- ts(Rice$ExportQuantity, 
                        start = min(Rice$Year), 
                        end = max(Rice$Year), 
                        frequency = 1)

RealExchangeRate_TS <- ts(Rice$RealVNDPerUSD, 
                          start = min(Rice$Year), 
                          end = max(Rice$Year), 
                          frequency = 1)

RealGDP_TS <- ts(Rice$RealGDP, 
                      start = min(Rice$Year), 
                      end = max(Rice$Year), 
                      frequency = 1)
```

### Examine stationarity of each time series

```{r}
library(tseries)

adf.test(ExportQuantity_TS)

adf.test(RealExchangeRate_TS)

adf.test(RealGDP_TS)
```

### Transform each time series into a stationary time series through differencing

```{r}
ExportQuantity_Diff <- diff(ExportQuantity_TS, differences = 1)

ExportQuantity_ADF <- adf.test(ExportQuantity_Diff)

ExportQuantity_ADF
```

```{r}
RealGDP_Diff <- diff(RealGDP_TS, differences = 2)

RealGDP_ADF <- adf.test(RealGDP_Diff)

RealGDP_ADF
```

### Combine all the time series into one matrix

```{r}
Rice_Matrix <- cbind(ExportQuantity_Diff, RealExchangeRate_TS, RealGDP_Diff) 

Rice_Matrix <- na.omit(Rice_Matrix)

Rice_Matrix
```

### Determine the lag order

```{r}
library(vars)

Rice_VARselect <- VARselect(Rice_Matrix, lag.max = 3, type = "none")

Rice_VARselect
```

### Estimate the reduced-form VAR

```{r}
Rice_VAR <- VAR(Rice_Matrix, p = 2, type = "none")

summary(Rice_VAR)
```

### Test if Exchange Rate Granger-causes Export Quantity

```{r}
Causality_RER <- causality(Rice_VAR, cause = "RealExchangeRate_TS")
Causality_RER
```

```{r}
Causality_GDP <- causality(Rice_VAR, cause = "RealGDP_Diff")
Causality_GDP
```

### Identify structural shocks with SVAR CV

```{r}
library(svars)

Rice_SVAR <- id.cv(Rice_VAR, SB = c(1985, 1)) # Specify the structural break in 1985

summary(Rice_SVAR)
```

### Assess the IRFs

```{r}
Rice_IRF <- irf(Rice_SVAR, n.ahead = 10)

plot(Rice_IRF)
```

### Add confidence intervals around IRFs

```{r}
Rice_IRF_CI <- wild.boot(Rice_SVAR, nboot = 500)

plot(Rice_IRF_CI)
```

### Compute FEVD

```{r}
Rice_FEVD <- fevd(Rice_SVAR, n.ahead = 10)

plot(Rice_FEVD)
```

### Compute the historical decomposition

```{r}
Rice_HD <- hd(Rice_SVAR)

plot(Rice_HD)
```

### Conduct counterfactual analysis without the exchange rate shock

```{r}
Rice_CF <- cf(Rice_SVAR)

plot(Rice_CF)
```

# OUTPUT EXPORTS

*Plot of monthly exchange rates*
```{r}
ggsave("MonthlyExchangeRate.pdf",
       plot = Plot_ExchangeRate_Monthly,
       width = 8,
       height = 4)
```

*Plot of annual nominal exchange rates*

```{r}
ggsave("NominalExchangeRate.pdf",
       plot = Plot_NER,
       width = 8,
       height = 4)
```

*Plot of annual real exchange rates*

```{r}
ggsave("RealExchangeRate.pdf",
       plot = Plot_RER,
       width = 8,
       height = 4)
```

*Plot of annual rice exports*

```{r}
ggsave("RiceExport.pdf",
       plot = Plot_Export,
       width = 8,
       height = 4)
```

*Plot of IRF with CI*

```{r}
pdf("IRF.pdf", width = 16, height = 8)
layout(matrix(1:16, nrow = 4, byrow = TRUE))
par(mar = c(4, 4, 3, 1))
plot(Rice_IRF_CI) 
dev.off()
```

*Plot of counterfactual analysis*

```{r}
pdf("Counterfactual.pdf", width = 8, height = 6)
plot(Rice_CF)
dev.off()
```

*Plot of FEVD*

```{r}
pdf("FEVD.pdf", width = 8, height = 6)
plot(Rice_FEVD)
dev.off()
```

*Plot of historical decomposition*

```{r}
pdf("HistoricalDecomposition.pdf", width = 8, height = 6)
plot(Rice_HD)
dev.off()
```
